---
title: "p8105_hw3_lz2949"
author: "Longyi Zhao"
date: "2023-10-09"
output: github_document
---

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

```{r}
num_aisles = unique(instacart$aisle_id) # number of distinct aisles numbers
frequent_aisles = 
  instacart %>% 
  group_by(aisle) %>%
  summarize(total_item = sum(order_number)) |>
  arrange(desc(total_item)) # most items ordered from 
```

make a plot that shows the number of items ordered in each aisle, (only aisles ordered more than 10000 items)
```{r}
item_aisles = instacart %>%
  group_by (aisle_id) %>%
  summarize(total_item = sum(order_number)) |>
  filter(total_item >= 10000) |>
  ggplot(aes(x = aisle_id, y = total_item)) +
  geom_bar(stat = "identity") +
  labs(title = "Number of Items sold in each Aisle (>=10000)")
 
print(item_aisles) 
```

Make a table showing the three most popular items in "baking ingredients", "dog food care", and "packaged vegetable fruits"

```{r}
three_aisle = instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits"))

# filter product name and add the order_number together
pop_baking = three_aisle  %>%
  filter(aisle == "baking ingredients") %>%
  group_by (product_name) %>%
  summarize(total_item = sum(order_number))
most_baking = pop_baking %>%
  arrange(desc(total_item)) %>%
  slice (1) %>%
  mutate(aisle = "baking ingredients")

# perform similar action to other two aisle
pop_dog = three_aisle  %>%
  filter(aisle == "dog food care") %>%
  group_by (product_name) %>%
  summarize(total_item = sum(order_number))
most_dog= pop_dog %>%
  arrange(desc(total_item)) %>%
  slice (1) %>%
  mutate(aisle = "dog food care")

pop_veg = three_aisle  %>%
  filter(aisle == "packaged vegetables fruits") %>%
  group_by (product_name) %>%
  summarize(total_item = sum(order_number))
most_veg= pop_veg %>%
  arrange(desc(total_item)) %>%
  slice (1) %>%
  mutate(aisle = "packaged vegetables fruits")


# arragne most_baking/dog/veg to a table 
combined_most = bind_rows(most_baking, most_dog, most_veg)
kable(combined_most, 
      caption = "Most popular items in three aisles") 
```

Make a table showing the mean hour of the day at which Pink lady apples and coffee ice cream are orderd on each day of the week. 

```{r}
app_cream = 
  instacart |>
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") |>
  group_by(order_dow, product_name) |>
  summarize(mean_time = mean(order_hour_of_day)) |>
  pivot_wider(names_from = product_name,
              values_from = mean_time) 
# app_cream = app_cream[, -1]
kable(app_cream, 
      caption = "Mean hour of the day ordered for Pink Lady Apples and Coffee Ice Cream")
```


## Problem 2
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Data cleaning 
```{r}
brfss_df = 
  brfss_smart2010 |>
  janitor::clean_names() |>
  filter(topic == "Overall Health") |>
  filter(response == "Excellent" | response =="Very good" |response =="Good" |response =="Fair" |response =="Poor") 

order_c = c("Poor", "Fair", "Good", "Very good", "Excellent")
ordered_brfss = brfss_df[order(factor(brfss_df$response, levels = order_c)),]
```

In 2002/2010, which states were observed at 7 or more locations?
```{r}
# year 2002
states_two = ordered_brfss |>
  filter(year == "2002") |>
  group_by(locationabbr) |>
  summarize(unique_locations = n_distinct(locationdesc)) 
states_two = states_two |>
  filter(unique_locations >= 7)

# year 2010
states_ten = ordered_brfss |>
  filter(year == "2010") |>
  group_by(locationabbr) |>
  summarize(unique_locations = n_distinct(locationdesc)) 
states_ten = states_ten |>
  filter(unique_locations >= 7)
```

Construct a dataset that is limited to Excellent responses and contains, year, state, and a variable averages the data_value accross locations within a state. Make a spaghetti plot. 

```{r}
exc_df = ordered_brfss |>
  filter(response == "Excellent") |>
  group_by(locationabbr, year) |>
  mutate(ave_data = mean(data_value)) |>
  select(year, locationabbr, ave_data) 

ggplot(exc_df, aes(x = year, 
                  y = ave_data,
                  group = locationabbr, 
                  color = locationabbr)) +
  geom_line() +
  labs(
    title = "Average Data Value Over Time by State (Excellent Responses)",
    x = "Year",
    y = "Average Data Value"
  )
```

Make a two-panel plot showing, for 2006/2010, distribution of data_value for responses (poor to excellent) among locations in NY state. 

```{r}
ny_df = ordered_brfss |>
  filter(locationabbr == "NY", year %in% c("2006", "2010")) 

ggplot(ny_df, aes(x = year, fill = response)) + geom_bar() +
  labs(
    title = "Distribution or Response by Year", 
    x = "Year", 
    y = "Count"
  )

# among locations??

```

## Problem 3

Load, tidy, merge, and organize the datasets: include all originally observed variables, exclude participants <21, exclude missing demographic data, encode data with reasonable variable classes. 

```{r}
nhanes_df = 
  read_csv("./data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  filter(age>=21) |>
  drop_na() |>
  mutate( 
    sex = case_match(
      sex,
      1 ~ "male", 
      2 ~ "female"
    ), 
    education=case_match(education, 
                           1~ "Less than high school", 
                         2 ~ "High school equivalent", 
                         3 ~ "More than high school"))

# drop na??
nhanes_accel = 
  read_csv("./data/nhanes_accel.csv") |>
  janitor::clean_names()

m_nhanes = inner_join(nhanes_df, nhanes_accel, by = "seqn")
```

Create a reader-friendly table for the number of men and women in each education category. 
```{r}
education_table = m_nhanes %>%
  group_by(education, sex) %>%
  summarize(count = n()) |>
  pivot_wider(names_from =  sex, 
              values_from = count)

kable(education_table, caption = "Sex and corresponding education level")
```

Create a visualization of the age distributions for men and women in each education category. 
```{r}
ggplot(education_table, aes(x= education))+
  geom_bar(aes(y = female, fill = "female"), position = "dodge", stat = "identity", alpha =0.5 )+
  geom_bar(aes(y= male, fill = "male"), position = "dodge", stat = "identity", alpha = 0.5) +
  labs(title = "Distribution of male and female in levels of education", 
       x = "education", 
       y = "number of individuals") 
```

Using tidies dataset, aggregate across minutes to create a total activity variable for each participant. Plot total activities (y) against age (x). Compare men to women and have separater panels for each education level. Include trend line or a smooth to illustrate differences. 

```{r}
sum_nhanes = m_nhanes |>
  rowwise() |>
  mutate(sum_acti = sum(c(min1:min1440))) |>
  select(-6:-1445)

ggplot(sum_nhanes, aes(x = age, y = sum_acti, color = sex)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) +
  facet_wrap(~ education) +
  labs(
    title = "Total Activity vs. Age by Education Level",
    x = "Age",
    y = "Total Activity"
  ) 
```





